% This file is to be included by LaTeX by the various WIP TeX files.
%
% Standard disclaimer, copyright, etc. notices. They go on inside cover page
% ------------------------------------------------------------------------
%  History:
%   pjt 
%   mjs 02apr91 Changed "Miriad" to MIRIAD, altered "Restrictions" section
%    jm 12jun91 Changed from MIRIAD to WIP
% ------------------------------------------------------------------------
\section*{\wip\ Description}

\htmladdnormallink{\wip\ is a graphics software package}{http://bima.astro.umd.edu/wip/wip.html}
available (at least) on Sun Workstations,
VAX/VMS systems, Linux machines, Cray/UNICOS, and CONVEX systems.
\wip\ is intended to be used to generate high quality graphics
(using the PGPLOT graphics library) with a minimum of effort.

\section*{Restrictions on Use}

\wip\ is copyright by the Berkeley-Illinois-Maryland Association (BIMA) Project.
All portions of \wip, including source code, may be used
subject only to the following restrictions:
\begin{enumerate}
  \item Unmodified \wip\ may be distributed freely on-site and
        may allow unrestricted access worldwide;
  \item Modified \wip\ must clearly note its non-standard nature and that
        portion derived from \wip\ can not be claimed as proprietary;
  \item Distribution of \wip\ requires formal notification to the
        \wip\ contact listed below;
  \item Receipt of \wip\ binds all recipients
        to the above restrictions and to all restrictions on portions of
        \wip\ not developed by BIMA (especially PGPLOT).
\end{enumerate}

\section*{Disclaimer}

We give no warranty, expressed or implied, for the software and/or
documentation provided, including, without limitation, warranty
of merchantability and warranty of fitness for a particular purpose.

\section*{Trademark Acknowledgments}

UNIX is a registered trademark of AT\&T.
CRAY and UNICOS are registered trademarks of Cray Research Inc.
Sun is a registered trademark and Sun Workstation a trademark of Sun
Microsystems Inc.
VAX and VMS are registered trademarks of Digital Equipment Corporation.
PGPLOT is copyright by the California Institute of Technology (all
rights reserved).

\section*{WIP Contact}

Please address all communication concerning \wip\ and this manual to: \\
\begin{flushleft}
\htmladdnormallink{\wip}{http://bima.astro.umd.edu/wip/wip.html}\\
\htmladdnormallink{James A. Morgan}{http://bima.astro.umd.edu/wip/author.html}\\
Astronomy Program \\
University of Maryland \\
College Park, MD  20742 \\
INTERNET:  morgan@astro.umd.edu \\
WWW:  http://bima.astro.umd.edu/wip/
\end{flushleft}
